rhdx
=====

```{r echo=FALSE}
library("knitr")
hook_output <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(output = function(x, options) {
   lines <- options$output.lines
   if (is.null(lines)) {
     return(hook_output(x, options))  # pass to default hook
   }
   x <- unlist(strsplit(x, "\n"))
   more <- "..."
   if (length(lines)==1) {        # first n lines
     if (length(x) > lines) {
       # truncate the output, but add ....
       x <- c(head(x, lines), more)
     }
   } else {
     x <- c(if (abs(lines[1])>1) more else NULL,
            x[lines],
            if (length(x)>lines[abs(length(lines))]) more else NULL
           )
   }
   # paste these lines together
   x <- paste(c(x, ""), collapse = "\n")
   hook_output(x, options)
 })

knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  eval = FALSE,
  warning = FALSE,
  message = FALSE
)
```

[![GitLab CI Build Status](https://gitlab.com/dickoa/rhdx/badges/master/build.svg)](https://gitlab.com/dickoa/rhdx/pipelines) 
[![AppVeyror Build status](https://ci.appveyor.com/api/projects/status/qytbcx7vjq0t9ao5/branch/master?svg=true)](https://ci.appveyor.com/project/dickoa/rhdx) 
[![Codecov Code Coverage](https://codecov.io/gl/dickoa/rhdx/branch/master/graph/badge.svg)](https://codecov.io/gl/dickoa/rhdx)
[![](http://www.r-pkg.org/badges/version/rhdx)](http://www.r-pkg.org/pkg/rhdx)
[![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/rhdx)](http://www.r-pkg.org/pkg/rhdx)
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)

`rhdx` is an R client for the Humanitarian Exchange Data platform.

## Installation

Development version

```{r}
install.packages("devtools")
devtools::install_git("https://gitlab.com/dickoa/rhdx")
```
## rhdx tutorial

```{r}
library("rhdx")
```

```{r}
Configuration$create(hdx_site = "test")
Configuration$read()
## <HDX Configuration> 
##   HDX site: test
##   HDX site url: https://test-data.humdata.org/
##   HDX API key: 
```

```{r}
datasets <- Dataset$search_in_hdx("ACLED", rows = 1)
datasets
## [[1]]
## <HDX Dataset> ac9f19f0-132e-46e6-9a9c-7d29dca8a469 
##   Title: ACLED Conflict Data for Algeria
##   Name: acled-conflict-data-for-algeria
##   Date: 01/01/1997-12/31/2015
##   Tags (up to 5): conflict, political violence, protests, war
##   Locations (up to 5): dza
##   Resources (up to 5): Algeria.xlsx
```

```{r}
resources <- datasets[[1]]$get_resources()
resources
## [[1]]
## <HDX Resource> 87ce2238-9049-4fb7-aa53-3c9d44b6183b 
##   Name: Algeria.xlsx
##   Description: 
##   Size: 
##   Format: XLSX
```

```{r}
resources[[1]]$download()
## trying URL 'http://www.acleddata.com/wp-content/uploads/2016/01/Eritrea.xlsx'
## Content type 'application/vnd.openxmlformats-officedocument.spreadsheetml.sheet' length 69983 bytes (68 KB)
## ==================================================
## downloaded 68 KB
```


## rhdx package API

+ Configuration - `create()` - `setup` - `read`
+ Dataset - `read_from_hdx` - `search_in_hdx`
+ Resource - `download` - `read_from_hdx` - `search_in_hdx`

## Future dev
### Develop all `create_in_hdx` methods
### Add tidy tools to easily get data
### Rstudio Add-in to browse data interactively

## Meta

* Please [report any issues or bugs](https://gitlab.dickoa/rhdx/issues).
* License: MIT
* Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.
