rhdx
=====

```{r echo=FALSE}
library("knitr")
hook_output <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(output = function(x, options) {
   lines <- options$output.lines
   if (is.null(lines)) {
     return(hook_output(x, options))  # pass to default hook
   }
   x <- unlist(strsplit(x, "\n"))
   more <- "..."
   if (length(lines)==1) {        # first n lines
     if (length(x) > lines) {
       # truncate the output, but add ....
       x <- c(head(x, lines), more)
     }
   } else {
     x <- c(if (abs(lines[1])>1) more else NULL,
            x[lines],
            if (length(x)>lines[abs(length(lines))]) more else NULL
           )
   }
   # paste these lines together
   x <- paste(c(x, ""), collapse = "\n")
   hook_output(x, options)
 })

knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  eval = FALSE,
  warning = FALSE,
  message = FALSE
)
```

[![GitLab CI Build Status](https://gitlab.com/dickoa/rhdx/badges/master/build.svg)](https://gitlab.com/dickoa/rhdx/pipelines) 
[![AppVeyror Build status](https://ci.appveyor.com/api/projects/status/qytbcx7vjq0t9ao5/branch/master?svg=true)](https://ci.appveyor.com/project/dickoa/rhdx) 
[![Codecov Code Coverage](https://codecov.io/gl/dickoa/rhdx/branch/master/graph/badge.svg)](https://codecov.io/gl/dickoa/rhdx)
[![](http://www.r-pkg.org/badges/version/rhdx)](http://www.r-pkg.org/pkg/rhdx)
[![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/rhdx)](http://www.r-pkg.org/pkg/rhdx)
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)

`rhdx` is an R client for the Humanitarian Exchange Data platform.

## Installation

Development version

```{r eval = FALSE}
install.packages("devtools")
devtools::install_git("https://gitlab.com/dickoa/rhdx")
```

```{r, eval = FALSE}
library('rhdx')
```


Note: the default base CKAN URL is set to
[http://data.techno-science.ca/](http://data.techno-science.ca/).
Functions requiring write permissions in CKAN additionally require a privileged
CKAN API key.
You can change this using `rhdx_setup()`, or change the URL using the `url`
parameter in each function call.
To set one or both, run:

```{r, eval = FALSE}
Configuration$create(prod = "demo") # restores default CKAN url to http://data.techno-science.ca/
```

```{r, eval = FALSE}
rhdx_setup(prod = "demo") # restores default CKAN url to http://data.techno-science.ca/
```


## rhdx package API
There are a suite of CKAN things (package, resource, etc.) that each have a set of functions in this package. The functions for each CKAN thing have an S3 class that is returned from most functions, and can be passed to most other functions (this also facilitates piping). The following is a list of the function groups for certain CKAN things, with the prefix for the functions that work with that thing, and the name of the S3 class:

+ Configuration - `create()` - `setup` - `read`
+ Dataset - `read_from_hdx` - `search_in_hdx`
+ Resource - `download` - `get_dataset` - `search_in_hdx`

The R6 class objects all look very similar; for example:

```r
<HDX Resource> 8abc92ad-7379-4fb8-bba0-549f38a26ddb
  Name: Data From Digital Portal
  Description:
  Size:
  Format: CSV
```

All classes state the type of object, have the ID to the right of the type, then have a varying set of key-value fields deemed important. This printed object is just a summary of an R list, so you can index to specific values (e.g., `result$description`). If you feel there are important fields left out of these printed summaries, let us know.

> note: Many examples are given in brief for readme brevity

## Packages

List packages

```{r}
Dataset$search_in_hdx(query = "Somalia IDP", rows = 5L)
```

### The National Geothermal Data System

Website: [http://geothermaldata.org/](http://geothermaldata.org/)

```{r}
rhdx_setup(hdx_site = "prod")
x <- Dataset$search_in_hdx(q = '*:*', rows = 1)
x$results
```

## Future dev
### Rstudio Add-in to browse data interactively

## Meta

* Please [report any issues or bugs](https://gitlab.dickoa/rhdx/issues).
* License: MIT
* Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.
