---
title: "R client to the Humanitarian Data Exchange"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tidy data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Humanitarian Data Exchange platform

The Humanitarian Data Exchange is the UN OCHA platform for share 

Development version from GitLab

```{r, eval = FALSE}
devtools::install_github("ropensci/crul")
devtools::install_git("https://gitlab.com/dickoa/rhdx")
```


```{r, eval = FALSE}
library("rhdx")
set_rhdx_config(hdx_site = "prod", read_only = TRUE)
```

## Get data from HDX using an OOP approach
### Connect to a server

You need to use the `Configuration` class to select a server 

```{r, eval = FALSE}
conf <- Configuration$setup(hdx_site = "prod")
conf
## <HDX Configuration> 
##   HDX site: prod
##   HDX site url: https://data.humdata.org/
##   HDX API key: 
```

### Search datasets

Once we selected a server, we can now search from dataset using the `search_in_hdx` method from the `Dataset` class.
In this case we will limit just to two results.

```{r, eval = FALSE}
list_of_ds <- Dataset$search_in_hdx("displaced Nigeria", rows = 2)
list_of_ds
## [[1]]
## <HDX Dataset> 4fbc627d-ff64-4bf6-8a49-59904eae15bb 
##   Title: Nigeria - Internally displaced persons - IDPs
##   Name: idmc-idp-data-for-nigeria
##   Date: 01/01/2009-12/31/2016
##   Tags (up to 5): displacement, idmc, population
##   Locations (up to 5): nga
##   Resources (up to 5): displacement_data, conflict_data, disaster_data

## [[2]]
## <HDX Dataset> 4adf7874-ae01-46fd-a442-5fc6b3c9dff1 
##   Title: Nigeria Baseline Assessment Data [IOM DTM]
##   Name: nigeria-baseline-data-iom-dtm
##   Date: 01/31/2018
##   Tags (up to 5): adamawa, assessment, baseline-data, baseline-dtm, bauchi
##   Locations (up to 5): nga
##   Resources (up to 5): DTM Nigeria Baseline Assessment Round 21, DTM Nigeria Baseline Assessment Round 20, DTM Nigeria Baseline Assessment Round 19, DTM Nigeria Baseline Assessment Round 18, DTM Nigeria Baseline Assessment Round 17
```
### Choose the dataset you want to manipulate in R, in this case we will take the first one.

```{r, eval = FALSE}
ds <- list_of_ds[[1]]
ds
## <HDX Dataset> 4fbc627d-ff64-4bf6-8a49-59904eae15bb 
##   Title: Nigeria - Internally displaced persons - IDPs
##   Name: idmc-idp-data-for-nigeria
##   Date: 01/01/2009-12/31/2016
##   Tags (up to 5): displacement, idmc, population
##   Locations (up to 5): nga
##   Resources (up to 5): displacement_data, conflict_data, disaster_data
```

### List all resources in the dataset

```{r, eval = FALSE}
list_of_rs <- ds$get_resources()
list_of_rs
## [[1]]
## <HDX Resource> f57be018-116e-4dd9-a7ab-8002e7627f36 
##   Name: displacement_data
##   Description: Internally displaced persons - IDPs (new displacement associated with conflict and violence)
##   Size: 
##   Format: JSON

## [[2]]
## <HDX Resource> 6261856c-afb9-4746-b340-9cf531cbd38f 
##   Name: conflict_data
##   Description: Internally displaced persons - IDPs (people displaced by conflict and violence)
##   Size: 
##   Format: JSON

## [[3]]
## <HDX Resource> b8ff1f4b-105c-4a6c-bf54-a543a486ab7e 
##   Name: disaster_data
##   Description: Internally displaced persons - IDPs (new displacement associated with disasters)
##   Size: 
##   Format: JSON
```

### Choose a resource we need to download/read
We are looking for the displacement data, it's the first resource in our list `list_of_rs`

```{r, eval = FALSE}
idp_nga_rs <- list_of_rs[[1]]
idp_nga_df <- idp_nga_rs$read_session(simplify_json = TRUE, folder = tempdir())
idp_nga_df
## $results
##   iso iso3 geo_name year conflict_new_displacements
## 1  NG  NGA  Nigeria 2009                       5000
## 2  NG  NGA  Nigeria 2010                       5000
## 3  NG  NGA  Nigeria 2011                      65000
## 4  NG  NGA  Nigeria 2012                      63000
## 5  NG  NGA  Nigeria 2013                     471000
## 6  NG  NGA  Nigeria 2014                     975000
## 7  NG  NGA  Nigeria 2015                     737000
## 8  NG  NGA  Nigeria 2016                     501000
##   disaster_new_displacements conflict_stock_displacement
## 1                     140000                          NA
## 2                     560000                          NA
## 3                       6300                          NA
## 4                    6112000                          NA
## 5                     117000                     3300000
## 6                       3000                     1075000
## 7                     100000                     2096000
## 8                      78000                     1955000

## $lookups
## named list()

## $errors
## list()

## $success
## [1] TRUE

## $params
## [1] "/api/displacement_data?iso3=NGA&ci=HDX00AKEYJUl17"

## $total
## [1] 8

## $limit
## [1] 0

## $offset
## [1] 0
```
